Titulo

### 2015-02-25 RotacionOutFlows/Paper/ms.tex

\shortauthors{Remolina-Gutierrez, Forero-Romero \& Garavito-Camargo}

\author{ Maria Camila Remolina-Gutierrez, Jaime E. Forero-Romero \& Juan N. Garavito-Camargo}
\affil{Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia}
\email{mc.remolina197@uniandes.edu.co}
\email{je.forero@uniandes.edu.co}
\email{jn.garavito57@uniandes.edu.co}

@@

This paper is structured as follows. In \S \ref{sec:theo} we explain in detail the model of rotation
and outflow that we use as well of our joint model (Rotation \&
Outflow). In \S \ref{sec:results} we present  the results of our
model. In \S \ref{sec:discussion} we compare our results with  recent
observations of LAEs with special attention of those morphologies that
present features of galaxy rotation and outflows. In the latest
section we present our  conclutions.

%LAEs are young and far away galaxies with most of its matter composed by Hydrogen atoms, being the rest heavier elements, like Iron, produced by growing stars burning their material. Then, this spectral line tells very important information about the galaxy, most of it actually, which creates a motivation to find a way to model the \lya line according to certain free parameters that vary from a galaxy to another and that can be obtained from observations. \\

%Our purpose in this paper is then to make a theoretical model, based on radiative transfer simulations, that use MonteCarlo
%computational methods to emulate how the line profile is going to come out of the galaxy if it is rotating and has a surrounding
%outflow, both of this phenomena characterized by key parameters that are going to be described further on the paper. \\

%As it has been described, the main purpose of this project is to mix a rotating galaxy with an external outflow surrounding it and analyze the resulting spectrum. The rotation model for the LAE consists on a sphere with an homogeneous mixture of dust and hydrogen at a constant temperature undergoing a solid-body rotation. The spectrum then depends on the maximum velocity of the sphere, the neutral hydrogen optical depths and the viewing angle. What happens is that photons are emitted from a central distribution of gas and escape the sphere after a radiative transfer process that ends at the border.\\

%In this model, rotation does not induce any spatial anisotropy in the integrated line flux, the escape fraction or the average
%number of scatterings, which led to the creation of an analytic approximation for the galaxy spectrum. This mathematical expression
%helps to minimize the computational costs and lets try with several parameters without spending great amounts of time running the code. \\

%After this photons escape the galaxy to empty space, they encounter a thin spherical shell of matter at a certain radius from the
%center. That material is the outflow of the LAE, created mainly by starbursts which composition is characterized with the parameters of Hydrogen column density and metallicity. Besides its components, the galactic outflow is also defined by an expanding constant velocity that affects the redshift of the photons.\\

%As there is a lot of empty space inside the shell, it acts as a reflector of photons. This means, that when one wants to comes out it is most likely to bounce inside the sphere several times before finally escaping. However it is not likely that the photon comes back again inside the galaxy, which helps that the first process does not have to be repeated, and both stages can be treated as consecutive, not simultaneous. \\

@@

We use the outflow model stated in \citep{Orsi12}. 
In their model the outflow consists of an homogeneous, expanding,
isothermal spherical shell surrounding a central galaxy.   
This model is completly described by three sets ot paremeters that
define the geometry, kinematics and thermodinamics of the model.
The geometry is described by the shell's inner radius $R_{in}$ and outer
radius $R_{out}$, which are related to each other by the expression
$R_{in} = f_{th}R_{out}$, with $f_{th}=0.9$ as the fiducial value. 
The temperature of the medium is assumed constant
and equal to $T=10^4 K$ and it is expanding at a constant velocity
\vel. This outflow is composed by dust and gas uniformly mixed
described by the column density \nh and the metallicity $Z$ as
follows:

\begin{equation}
\label{eq:nh}
n_H = \frac{X_H M_{shell}}{4\pi m_H {R^2}_{out}}
\end{equation}

Where $M_{shell}$ is the mass of the outflow, $m_H$ is the mass of the
hydrogen atom and $X_H=0.74$ is the fraction of hydrogen in the cold
gas.

\begin{equation}
\label{eq:z}
Z = f_Z \langle Z_{cold} \rangle
\end{equation}

Where $\langle Z_{cold} \rangle$ is the average metallicity of the
cold gas and it is weighted by the product of the mass of cold gas and
the \lya luminosity of the disc and bulge $f_Z$.

We decided to choose this model for two main reasons. First, when we
analyze observed spectra we can se that the line is spread around the
place where line should be when already taking into account the
redshift due to the distance to the LAE. This implies that the photons
before escaping the composed system suffer a change in frequency
themselves, which we explain by making them rebound over the inner
surface of the shell such as some of them suffer blueshift and other
redshift, that even small, is significant and observed.

The second reason is because as the space between the shell and the
galaxy is said to be void, the photons that escape from it suffer the
scattering process at a far distance from the center, in a way such as
if they rebound the chances that it happens in the normal direction to
the surface are lower than in any other model. This implies that the
photons are less likely to re-enter the galaxy so there is no need to
start the diffusion inside again, allowing us to divide the spectrum
creation process into 2 independent events, saving computational time
and simplifying the model to its best.

>>>>

\shortauthors{Remolina-Gutierrez et al.}

\author{Maria Camila Remolina-Gutierrez, Jaime E. Forero-Romero \&
  Juan N. Garavito-Camargo} 
\affil{Departamento de F\'{i}sica, Universidad de los Andes, Cra. 1
No. 18A-10, Edificio Ip, Bogot\'a, Colombia}
\author{Julian Mej\'ia}
\affil{Chile}
\author{Alvaro Orsi}
\affil{Spain}
\email{mc.remolina197@uniandes.edu.co}
\email{je.forero@uniandes.edu.co}
\email{jn.garavito57@uniandes.edu.co}

@@

This paper is structured as follows. In \S \ref{sec:theo} we explain
in detail the model of rotation and outflow that we use as well of our
joint model (Rotation \& Outflow). In \S \ref{sec:results} we present
the results of our model. In \S \ref{sec:discussion} we compare our
results with  recent observations of LAEs with special attention of
those morphologies that present features of galaxy rotation and
outflows. In the latest section we present our  conclusions.

%LAEs are young and far away galaxies with most of its matter composed
%by Hydrogen atoms, being the rest heavier elements, like Iron,
%produced by growing stars burning their material. Then, this spectral
%line tells very important information about the galaxy, most of it
%actually, which creates a motivation to find a way to model the \lya
%line according to certain free parameters that vary from a galaxy to
%another and that can be obtained from observations. \\

%Our purpose in this paper is then to make a theoretical model, based
%on radiative transfer simulations, that use MonteCarlo 
%computational methods to emulate how the line profile is going to
%come out of the galaxy if it is rotating and has a surrounding 
%outflow, both of this phenomena characterized by key parameters that
%are going to be described further on the paper. \\

%As it has been described, the main purpose of this project is to mix
%a rotating galaxy with an external outflow surrounding it and analyze
%the resulting spectrum. The rotation model for the LAE consists on a
%sphere with an homogeneous mixture of dust and hydrogen at a constant
%temperature undergoing a solid-body rotation. The spectrum then
%depends on the maximum velocity of the sphere, the neutral hydrogen
%optical depths and the viewing angle. What happens is that photons
%are emitted from a central distribution of gas and escape the sphere
%after a radiative transfer process that ends at the border.\\

%In this model, rotation does not induce any spatial anisotropy in the
%integrated line flux, the escape fraction or the average 
%number of scatterings, which led to the creation of an analytic
%approximation for the galaxy spectrum. This mathematical expression 
%helps to minimize the computational costs and lets try with several
%parameters without spending great amounts of time running the
%code. \\

%After this photons escape the galaxy to empty space, they encounter a thin spherical shell of matter at a certain radius from the
%center. That material is the outflow of the LAE, created mainly by
%starbursts which composition is characterized with the parameters of
%Hydrogen column density and metallicity. Besides its components, the
%galactic outflow is also defined by an expanding constant velocity
%that affects the redshift of the photons.\\

%As there is a lot of empty space inside the shell, it acts as a
%reflector of photons. This means, that when one wants to comes out it
%is most likely to bounce inside the sphere several times before
%finally escaping. However it is not likely that the photon comes back
%again inside the galaxy, which helps that the first process does not
%have to be repeated, and both stages can be treated as consecutive,
%not simultaneous. \\

@@

We use an outflow model that follows the characteristics put described
in \citep{Verhamme06} with the code presented in\citep{Orsi12}.  
The outflow consists of an isothermal, spherical flow expanding at
constant velocity \vel.  
The outflow is empty inside the shell's inner radius $R_{in}$ and
reaches out to an external radius $R_{out}$.
The relationship between these two radii is parameterized by $R_{in} =
f_{th}R_{out}$, with $f_{th}=0.9$ as the fiducial value.  
The temperature of the medium is assumed constant and equal to $T=10^4
K$ which sets the velocity dispersion of Maxwell-Boltzmann
distributions to $v_{\rm th}=12.84$\kms.

The gas has an homogeneous Hydrogen number density inside the flow. 
The total mass inside the shell is parameterized by its column density
\begin{equation}
\label{eq:nh}
N_H = \frac{X_H M_{shell}}{4\pi m_H {R^2}_{out}},
\end{equation}
%
where $M_{shell}$ is the outflow mass, $m_H$ is the mass of the
hydrogen atom and $X_H=0.74$ is the fraction of hydrogen in the cold
gas.

The outflow also includes dust homogeneously mixed with the gas. 
The dust optical depth $\tau_d$ is parameterized by the metallicity of
the cold gas $\langle Z_{cold} \rangle$: 
\begin{equation}
\label{eq:z}
\tau_{d} ???? \langle Z_{cold} \rangle
\end{equation}

%We choose this model because it h
%for two main reasons. First, when we
%analyze observed spectra we can se that the line is spread around the
%place where line should be when already taking into account the
%redshift due to the distance to the LAE. This implies that the photons
%before escaping the composed system suffer a change in frequency
%themselves, which we explain by making them rebound over the inner
%surface of the shell such as some of them suffer blueshift and other
%redshift, that even small, is significant and observed.

%The second reason is because as the space between the shell and the
%galaxy is said to be void, the photons that escape from it suffer the
%scattering process at a far distance from the center, in a way such as
%if they rebound the chances that it happens in the normal direction to
%the surface are lower than in any other model. This implies that the
%photons are less likely to re-enter the galaxy so there is no need to
%start the diffusion inside again, allowing us to divide the spectrum
%creation process into 2 independent events, saving computational time
%and simplifying the model to its best.
